Robustly decorrelating errors with mixed quantum gates

### 2019-08-05 decorrelation/mixed_unitaries.tex
bye bye norms

Campbell \cite{Campbell2017} and Hastings \cite{1612.01011} independently considered this problem using gates constructed with the Solovey-Kitaev algorithm, for which many approximate gate compilations are possible. Campbell showed that, if the error generators of some ensemble of gate compilations form a convex set containing the origin, then one can construct a MQG with quadratically suppressed diamond distance to the target. Explicitly, the weights are chosen such that the error generator is canceled to first order. Using (\ref{eq:effective_error}) and (\ref{eq:generator}), the effective error map for a MQG in terms of the component error generators is:
\begin{align}
	\label{eq:generators_and_effective_error}
	\mathcal{E}_{\rm eff} \simeq \left(\mathcal{I}_d + \sum_i w_i \mathcal{L}_i + \frac{1}{2} \sum_i w_i \mathcal{L}_i^2\right)
\end{align}
If, as Campbell required, the origin is in the convex hull of the generators (see Fig.~\ref{fig:vectorspace}), then there exists a choice for the weights, $w_i^*$, such that $\sum_i w^*=1$ and:
\begin{align}
	\label{eq:cancel}
	&||\sum_i w_i^* \mathcal{L}_i||= 0
	% \mathrm{with}\;&\sum_i w_i = 1
\end{align}
where $||\cdot||$ is the Frobenius norm. \kcy{why the norm, why not just $\sum_i w_i L_i = 0$}.

@@

One reason to construct an MQG is to improve the worst-case performance of a quantum gate. The worst-case performance is bounded by the diamond distance to the target, which would constitute a natural optimization target.  The diamond distance, however, is a non-linear function that in general requires a convex optimization to compute\cite{watrous2018theory}. This can dramatically slow down iterative optimizers, and so minimizing the diamond distance is a computationally expensive optimization target. However, if the errors are small enough, we can consider a linearized problem, minimizing the diamond distance to first-order in the effective error generator. As discussed in \cite{Campbell2017}, a sufficient condition to minimize the diamond distance of an MQG with error generators $\{\mathcal{L}_i\}$ to first order is:
\begin{equation}\label{eq:campbell-condition}
||\sum \omega_i \mathcal{L}_i ||= 0
\end{equation}
\kcy{again with the norm}
In \cite{Campbell2017} Campbell presents an algorithm that, given an oracle that produces approximate unitary operations, will generate controls and weightings to find an MQG with this property. Alternatively, one might consider the following problem: \textit{Given a fixed collection of controls, how can I produce a weighting with minimal diamond norm?} Such a situation might arise if there is a natural family of controls that implement a desired gate, or if, as we consider later (Section \ref{sec:numerical_results}), we are given a collection of precomputed gate implementations. Given such an MQC we can consider the $n^2\times m$ matrix whose columns are the vectorized error generators:
\begin{equation}\label{eq:vectorized_generators}
	\bold{L} = \left(\begin{array}{cccc}
		\mathcal{L}_{1_{1,1}} & \mathcal{L}_{2_{1,1}} & \ldots   \\ 
		\vdots\ & \ddots &    \\
		\mathcal{L}_{1_{n,n}} &  &  \mathcal{L}_{m_{n,n}} \\ 
	\end{array} 	
	\right)
\end{equation}
To satisfy (\ref{eq:campbell-condition}), we can write the sum as a matrix product and solve the following optimization problem, which will be zero if and only if the sum is zero:

@@

% ==============================================================================
% Section: Robustness
% ==============================================================================
\subsection{Additional Constraints}
\subsubsection{Adding robustness} % (fold)
\label{sub:adding_robustness}
While mixed quantum channels offer significant improvements to gate performance, they fail to take into account the reality that most control electronics experience drift over time scales relevant to QIP performance. Because of this drift, the quality of the MQG will degrade. Thus, we would like to design MQGs that are robust to this drift. To enforce robustness, we can consider the generators in (\ref{eq:campbell-condition}) to be functions of a control vector $\vec{\delta}$, and instead of requiring that the error generators average to zero, we can impose a similar condition on the higher-order derivatives with respect to parameters that may drift. Specifically, we are interested in the $n^{th}$ order derivatives:
\begin{equation}
\label{eq:derivative}
D^n_i = \frac{1}{n!}\frac{\partial^{n}}{\partial\delta_{j_1}\ldots\partial\delta_{j_n}}\mathcal{L}_i(\vec{\delta})|_{\vec{\delta}=\vec{0}}
\end{equation}
If the dimension of $\vec{\delta}$ is $d$, the indices $j_0, \ldots, j_n$ take on values in $1, \ldots, d$ and this matrix has $d^n$ entries. 
We say that an MQG is robust to order $\ell$ (an $\ell$MQG) if for all $1 \leq k \leq \ell$:
\begin{equation}\label{eq:MQG}
\begin{gathered}
||\sum_i\omega_i(\sum_{n=0}^k D^n_i)^n|| = 0\\
\end{gathered}
\end{equation}
In particular, we see that a 0MQG satisfies (\ref{eq:campbell-condition}). More generally, these conditions imply that an $\ell$MQG is insensitive to $\ell^{th}$ order in drift in $\vec{\delta}$. To see this, we can rewrite the implementations of the target gate as:
\begin{equation}\label{eq:taylor}
\begin{gathered}
\tilde{\mathcal{G}_i}(d\vec{\delta}) = \exp(-i(\mathcal{L}_i(\vec{0}) + \frac{\partial}{\partial\delta_j}\mathcal{L}_i(\vec{0})d\delta_j\\ +  \frac{1}{2}\frac{\partial^2}{\partial\delta_j\partial\delta_k} \mathcal{L}_i(\vec{0})d\delta_j d\delta_k + \ldots))\mathcal{G}
\end{gathered}
\end{equation}
By Taylor expanding (\ref{eq:taylor}) in $d\vec{\delta}$, one can see that the the first $\ell$ derivatives of an $\ell$MQG will be zero. Furthermore if we can find controls such that $||D_i^n||\leq\epsilon$, we can approximate (\ref{eq:MQG}) as:
\begin{equation}\label{eq:MQG-relaxed}
\begin{gathered}
||\sum\omega_iD^n_i||+ \mathcal{O}(\epsilon^2) = 0\\
\end{gathered}
\end{equation}
This condition guarantees that errors will be suppressed to second order in $\epsilon$ for all derivatives up to order $\ell$. %A proof is included in the appendix that generalizes the Hastings-Campbell Mixing Lemma in \cite{Campbell2017}. Namely, Campbell showed that if $0\in $ Conv$[\{G_i(\vec{\delta})\}]$, where Conv is the convex hull of its arguments, then an $\vec{\omega}$ exists that quadratically decreases the diamond norm. We prove that $0\in $ Conv$[\{D_j^n(\vec{\delta})\}]$ implies there is an $\vec{\omega}$ exists that quadratically decreases the $\ell^{th}$-order sensitivity of the diamond norm of an $\ell$MQG.
As discussed in \cite{Campbell2017}, the geometric picture for 0MQGs is to consider the error generators as vectors, and to ensure zero is contained in their convex hull. Figure \ref{fig:vectorspace} generalizes this picture to 1MQGs and gives intuition for the conditions required to produce them. To generate $\ell$MQGs, we first define the vectorized derivative matrix ${\bold{D}^{\ell}}$ in a similar way to (\ref{eq:vectorized_generators}):
\begin{equation}
{\bold{D}^{\ell}} =  \left(\begin{array}{cccc}
		D^\ell_{1_{1,1}} & D^\ell_{2_{1,1}} & \ldots   \\ 
		\vdots\ & \ddots &    \\
		D^\ell_{1_{n,n}} &  &  D^\ell_{m_{n,n}} \\ 
	\end{array} 	
	\right)
\end{equation}

>>>>

Campbell \cite{Campbell2017} and Hastings \cite{1612.01011} independently considered this problem using gates constructed with the Solovey-Kitaev algorithm, for which many approximate gate compilations are possible. Campbell showed that, if the error generators of some ensemble of gate compilations form a convex set containing the origin, then one can construct a MQG with quadratically suppressed diamond distance to the target. Explicitly, the weights are chosen such that the error generator is canceled to first order. Using (\ref{eq:effective_error}) and (\ref{eq:generator}), the effective error map for a MQG in terms of the component error generators is:
\begin{align}
	\label{eq:generators_and_effective_error}
	\mathcal{E}_{\rm eff} \simeq \left(\mathcal{I}_d + \sum_i w_i \mathcal{L}_i + \frac{1}{2} \sum_i w_i \mathcal{L}_i^2\right)
\end{align}
If, as Campbell required, the origin is in the convex hull of the generators (see Fig.~\ref{fig:vectorspace}), then there exists a choice for the weights, $w_i^*$, such that $\sum_i w^*=1$ and:
\begin{align}
	\label{eq:cancel}
	&\sum_i w_i^* \mathcal{L}_i= 0
	% \mathrm{with}\;&\sum_i w_i = 1
\end{align}

@@

One reason to construct an MQG is to improve the worst-case performance of a quantum gate. The worst-case performance is bounded by the diamond distance to the target, which would constitute a natural optimization target.  The diamond distance, however, is a non-linear function that in general requires a convex optimization to compute\cite{watrous2018theory}. This can dramatically slow down iterative optimizers, and so minimizing the diamond distance is a computationally expensive optimization target. However, if the errors are small enough, we can consider a linearized problem, minimizing the diamond distance to first-order in the effective error generator. As discussed in \cite{Campbell2017}, a sufficient condition to minimize the diamond distance of an MQG with error generators $\{\mathcal{L}_i\}$ to first order is:
\begin{equation}\label{eq:campbell-condition}
\sum \omega_i \mathcal{L}_i = 0
\end{equation}
\kcy{again with the norm}
In \cite{Campbell2017} Campbell presents an algorithm that, given an oracle that produces approximate unitary operations, will generate controls and weightings to find an MQG with this property. Alternatively, one might consider the following problem: \textit{Given a fixed collection of controls, how can I produce a weighting with minimal diamond norm?} Such a situation might arise if there is a natural family of controls that implement a desired gate, or if, as we consider later (Section \ref{sec:numerical_results}), we are given a collection of precomputed gate implementations. Given such an MQC we can consider the $n^2\times m$ matrix whose columns are the vectorized error generators:
\begin{equation}\label{eq:vectorized_generators}
	\bold{L} = \left(\begin{array}{cccc}
		\mathcal{L}_{1_{1,1}} & \mathcal{L}_{2_{1,1}} & \ldots   \\ 
		\vdots\ & \ddots &    \\
		\mathcal{L}_{1_{n,n}} &  &  \mathcal{L}_{m_{n,n}} \\ 
	\end{array} 	
	\right)
\end{equation}
To satisfy (\ref{eq:campbell-condition}), we can write the sum as a matrix product and solve the following optimization problem, which will be zero if and only if the sum is zero:

@@

% ==============================================================================
% Section: Robustness
% ==============================================================================
\subsection{Additional Constraints}
\subsubsection{Adding robustness} % (fold)
\label{sub:adding_robustness}
While mixed quantum channels offer significant improvements to gate performance, they fail to take into account the reality that most control electronics experience drift over time scales relevant to QIP performance. Because of this drift, the quality of the MQG will degrade. Thus, we would like to design MQGs that are robust to this drift. To enforce robustness, we can consider the generators in (\ref{eq:campbell-condition}) to be functions of a control vector $\vec{\delta}$, and instead of requiring that the error generators average to zero, we can impose a similar condition on the higher-order derivatives with respect to parameters that may drift. Specifically, we are interested in the $n^{th}$ order derivatives:
\begin{equation}
\label{eq:derivative}
D^n_i = \frac{1}{n!}\frac{\partial^{n}}{\partial\delta_{j_1}\ldots\partial\delta_{j_n}}\mathcal{L}_i(\vec{\delta})|_{\vec{\delta}=\vec{0}}
\end{equation}
If the dimension of $\vec{\delta}$ is $d$, the indices $j_0, \ldots, j_n$ take on values in $1, \ldots, d$ and this matrix has $d^n$ entries. 
We say that an MQG is robust to order $\ell$ (an $\ell$MQG) if for all $1 \leq k \leq \ell$:
\begin{equation}\label{eq:MQG}
\begin{gathered}
\sum_i\omega_i(\sum_{n=0}^k D^n_i)^n = 0\\
\end{gathered}
\end{equation}
In particular, we see that a 0MQG satisfies (\ref{eq:campbell-condition}). More generally, these conditions imply that an $\ell$MQG is insensitive to $\ell^{th}$ order in drift in $\vec{\delta}$. To see this, we can rewrite the implementations of the target gate as:
\begin{equation}\label{eq:taylor}
\begin{gathered}
\tilde{\mathcal{G}_i}(d\vec{\delta}) = \exp(-i(\mathcal{L}_i(\vec{0}) + \frac{\partial}{\partial\delta_j}\mathcal{L}_i(\vec{0})d\delta_j\\ +  \frac{1}{2}\frac{\partial^2}{\partial\delta_j\partial\delta_k} \mathcal{L}_i(\vec{0})d\delta_j d\delta_k + \ldots))\mathcal{G}
\end{gathered}
\end{equation}
By Taylor expanding (\ref{eq:taylor}) in $d\vec{\delta}$, one can see that the the first $\ell$ derivatives of an $\ell$MQG will be zero. Furthermore if we can find controls such that $||D_i^n||\leq\alpha$, (\ref{eq:MQG}) is approximately satisfied in the sense that:
\begin{equation}\label{eq:MQG-relaxed}
\begin{gathered}
||\sum\omega_iD^n_i||+ \mathcal{O}(\alpha^2) = 0\\
\end{gathered}
\end{equation}
where $||\cdot||$ is the Frobenius norm. This condition guarantees, as before, that errors will be suppressed to second order in $\alpha$, but now for all derivatives up to order $\ell$. %A proof is included in the appendix that generalizes the Hastings-Campbell Mixing Lemma in \cite{Campbell2017}. Namely, Campbell showed that if $0\in $ Conv$[\{G_i(\vec{\delta})\}]$, where Conv is the convex hull of its arguments, then an $\vec{\omega}$ exists that quadratically decreases the diamond norm. We prove that $0\in $ Conv$[\{D_j^n(\vec{\delta})\}]$ implies there is an $\vec{\omega}$ exists that quadratically decreases the $\ell^{th}$-order sensitivity of the diamond norm of an $\ell$MQG.
As discussed in \cite{Campbell2017}, the geometric picture for 0MQGs is to consider the error generators as vectors, and to ensure zero is contained in their convex hull. Figure \ref{fig:vectorspace} generalizes this picture to 1MQGs and gives intuition for the conditions required to produce them. To generate $\ell$MQGs, we first define the vectorized derivative matrix ${\bold{D}^{\ell}}$ in a similar way to (\ref{eq:vectorized_generators}):
\begin{equation}
{\bold{D}^{\ell}} =  \left(\begin{array}{cccc}
		D^\ell_{1_{1,1}} & D^\ell_{2_{1,1}} & \ldots   \\ 
		\vdots\ & \ddots &    \\
		D^\ell_{1_{n,n}} &  &  D^\ell_{m_{n,n}} \\ 
	\end{array} 	
	\right)
\end{equation}
